---
title: "Lab 2 Assignment: Group 04"
format:
  html:
    embed-resources: true
editor: visual
---

### Group Exercise from Group 4 

- Edir Sebastian Vidal Castro (s243564), 
- Morten Orebo Holmström (s242223),
- Nithiyashri Jayashankar (s244356),
- Tilman Johannes Nelissen (s233043),
- Thea Rehm Rosholm (s225019)

### Downloading packages

(remove comment for the corresponding packages)

```{r}
#install.packages("tidyverse")
#install.packages("ggridges")
#install.packages("RColorBrewer")
#install.packages("curl")
```

### Loading packages

```{r}
library("tidyverse")
library("ggridges")
library("RColorBrewer")
library("curl")
```

### Reading in the data

Using the curl_download with arguments url = the webpage/URL, destfile = the name of the data file.

```{r}
target_url <- "https://github.com/ramhiser/datamicroarray/raw/master/data/gravier.RData" 

main_dir <- getwd()

if(file.exists("data/")==FALSE){
  dir.create(file.path(main_dir, "data/"))
}

output_file <- "data/gravier.RData" 

curl::curl_download(url = target_url,
                    destfile = output_file)
```

Loading output file and checking that the file is in our environment.

```{r}
load(output_file) 
ls()
```

### Data wrangling

```{r}
set.seed(676571)

cancer_data <- gravier |> 
  bind_cols() |> 
  rename(event_label = y) |>
  mutate(pt_id = 1:length(pluck(gravier, "y")),
         age = round(rnorm(length(pluck(gravier, "y")),mean=55,sd=10),1))
  
cancer_data$age_group <- cut(cancer_data$age, breaks=seq(10,100,by=10))

cancer_data <- cancer_data |>
  relocate(c(pt_id,age,age_group,pt_id,event_label))
```

We chose the probe g1A08 and use the package ggridges to show the distribution of probe fluorescence between patients with a good outcome and poor outcome.

```{r}
cancer_data |>
  ggplot(aes(x = g1A08, 
             y = event_label, fill = event_label)) +
  geom_density_ridges(alpha = 0.7) +
  theme_bw() +
  labs(y = "Outcome",
       x = "Fluoresence",
       title = "Fluorescence of probe g1A08 (log2 transformed) stratified on patient \noutcome") +
  theme(legend.position = "none") + 
  scale_fill_brewer(palette = "Set2")
```
