---
title: "Lab 4 Assignment Group 4"
format: html
editor: visual
execute: 
  warning: false
---

### Group Exercise from Group 4

-   Edir Sebastian Vidal Castro (s243564),
-   Morten Orebo Holmström (s242223),
-   Nithiyashri Jayashankar (s244356),
-   Tilman Johannes Nelissen (s233043),
-   Thea Rehm Rosholm (s225019)

### Loading packages

```{r}
#| message: false
library("tidyverse")
library("patchwork")
library("here")
```

### Loading data

We use the `here()` function to go to the root directory of our project.

```{r}
#diabetes_data <- read_csv(file = here("data/diabetes (1).csv"))
```

Or alternatively we can access the data through its URL.

```{r}
diabetes_data <- read_csv(file = "https://hbiostat.org/data/repo/diabetes.csv")
```

### T1

Change `height`, `weight`, `waist` and `hip` variables from imperial to metric system.

We need to use `mutate` to do this but first we need to know how to recalculate from imperial to metric. Conversion of inches to cm = inches x 2.54 Conversion of pounds to kg = pounds x 0.45359237

And we should round to one decimal using the `round` function.

```{r}
diabetes_data <- diabetes_data |> 
  mutate(m_height = height * 2.54,
         m_weight = weight * 0.45359237,
         m_waist = waist * 2.54,
         m_hip = hip * 2.54) |> 
           mutate(across(starts_with("m_"), 
                         ~round(., 1)))
```

Okay, so an important lesson here is that we can use the `mutate(across())` function to apply a subsequent function to mutate. So here we mutate across variables that `starts_with("m_")` which are our newly calculated variables, and then we apply the function `round` to these variables. As we use the `across` function we need to put the variables inside round as a dot i.e. `round(., digits = 1)` and we need to provide a \~ before the `round` function.

### Q17

Creating new variable `BMI` which is formula is as follows:

$$
BMI = \frac{weight_{kg}}{(height_{m})^2}
$$

So applying our variables it would be `m_weight/((m_height/100)^2)` for every observation.

```{r}
diabetes_data <- diabetes_data |> 
  mutate(BMI = m_weight/((m_height/100)^2))
```

### T7

Creating the variable `BFP` based on the equation:

$$
BFP = 1.39 \cdot BMI + 0.16 \cdot age - 10 \cdot gender - 9
$$

Where gender is 0 for female and 1 for male.

So essentially we can approach this problem by creating a new variable termed `gender_number` which gets 0 for female and 1 for male.

```{r}
diabetes_data <- diabetes_data |> 
  mutate(gender_number = case_when(gender == "female" ~ 0, 
                                   gender == "male" ~ 1))
```

And then we can construct our new variable based on the formula `mutate(BFP = 1.39 * BMI + 0.16 * age - 10 * gender_number - 9)` for every observation.

```{r}
diabetes_data <- diabetes_data |> 
  mutate(BFP = 1.39 * BMI + 0.16 * age - 10 * gender_number - 9)
```

### T8

Creating a `WHR` variable which is defined by the ratio among the waist and the hip: `mutate(WHR = m_waist/m_hip)`.

```{r}
diabetes_data <- diabetes_data |> 
  mutate(WHR = m_waist/m_hip)
```

### Q18 - GROUP ASSIGNMENT

What correlates better with BMI - WHR or BFP?

We can use the correlation coefficient for BMI vs. WHR and BFP respectively to find out which variable correlates the best with BMI. For this we can use the Pearson Correlation Coefficient (r) applied to the variables as follows:

$$
r = \frac{\sum (x_i - \bar{x})(y_i - \bar{y})}{\sqrt{\sum(x_i - \bar{x})^2 \sum(y_i - \bar{y})^2}}
$$

Where $y$ will be our variable BMI and $x$ either WHR or BFP. This computation can be done by applying the `cor()` function using the corresponding variables.

BMI correlation to WHR:

```{r}
cor(diabetes_data$BMI, diabetes_data$WHR, use = "complete.obs")
```

Correlation coefficient between DMI and WHR of 0.1 - a low correlation coefficient.

BMI correlation to BFP:

```{r}
cor(diabetes_data$BMI, diabetes_data$BFP, use = "complete.obs")
```

And for the BFP it is 0.89. Considerably higher correlation coefficient meaning, that BFP is better associated with BMI than WHR.

We use the geom_point to asses the association between BMI, WHR and BFP graphically and save the plot in the object `whr_plot`. We apply a regression line to show the relationship between BMI and WHR

```{r}
whr_plot <- diabetes_data |> 
  ggplot(aes(x = BMI, 
             y = WHR)) + 
  geom_point(alpha =0.6,
             fill = "steelblue", 
             shape = 21, 
             size = 2) +
  labs(title = "Correlation between body mass\n index and waist to hip ratio", 
       y = "Waist to hip ratio", 
       x = "Body mass index",
       caption = "Pearsson correlation coefficent = 0.10")  + 
  geom_smooth(method = "lm",
              se = F) + 
  theme_bw()

whr_plot
```

We use the geom_point to asses the association between BMI, WHR and BFP graphically and save the plot in the object `bfp_plot`. We apply a regression line to show the relationship between BMI and BFP.

```{r}
bfp_plot <- diabetes_data |> 
  ggplot(aes(x = BMI, 
             y = BFP)) + 
  geom_point(alpha =0.6,
             fill = "steelblue", 
             shape = 21, 
             size = 2) +
  labs(title = "Correlation between body mass\n index and body fat percentage", 
       y = "Body fat percentage", 
       x = "Body mass index", 
       caption = "Pearsson correlation coefficent = 0.89")  + 
  geom_smooth(method = "lm",
              se = F) + 
  theme_bw()

bfp_plot
```

We use that `patchwork` package merge the two plots into one nice plot.

```{r}
#| message: false 
whr_plot + bfp_plot
```
