---
title: "Lab 2 Assignment: Group 04"
format:
  html:
    embed-resources: true
editor: visual
---

### Group Exercise from Group 4 

- Edir Sebastian Vidal Castro (s243564), 
- Morten Orebo Holmström (s242223),
- Nithiyashri Jayashankar (s244356),
- Tilman Johannes Nelissen (s233043),
- Thea Rehm Rosholm (s225019)


### Loading packages
```{r}
library("tidyverse")
library("RColorBrewer")
library("here")
```

Loading the data. We use the `here` function, as we have our working directory in the scripts. We want to save within the `data/_raw` folder and also load the data from `here`. That is why we use the `here` function.

```{r}
#| echo: true
#| eval: false

base_url <- "http://userweb.eng.gla.ac.uk/umer.ijaz/bioinformatics/ecological/"
raw_data_path <- "data/_raw/"

SPE <- read_csv(file = str_c(base_url, "SPE_pitlatrine.csv"))

write_csv(x = SPE,
          file = here(str_c(raw_data_path, "SPE_pitlatrine.csv")))

ENV <- read_csv(file = str_c(base_url, "ENV_pitlatrine.csv"))

write_csv(x = ENV,
          file = here(str_c(raw_data_path, "ENV_pitlatrine.csv")))
```




```{r}
#| eval: true
#| echo: true

SPE |> 
  pivot_longer(cols = -Taxa,
               names_to = "Samples",
               values_to = "OTU_Count")  |> 
  full_join(ENV, by = "Samples") |> 
  mutate(site = case_when(str_detect(Samples, "^T") ~ "Tanzania",
                          str_detect(Samples, "^V") ~ "Vietnam")) |>  
  write_tsv(file = here("data/SPE_ENV.tsv"))
```

Loading the data
```{r}
SPE_ENV <- read_tsv(file = here("data/SPE_ENV.tsv"))
SPE_ENV
```





### Writing code for our figure
```{r}
SPE_ENV %>% 
  filter(Taxa == "Bacteroidia") %>% #Filtering for Bacteroidia only 
ggplot(mapping = aes(x = Prot, # Mapping to proteins and carbohydrates
                     y = Carbo)) + 
  geom_point(aes(size = OTU_Count, #Showing the OTU counts as the size of the geom_point
                 fill = site), #coloring the points according to their site
             shape = 21, #allows for a circle around the points
             alpha = 0.5, # adding some transparency due to data overlap
             color = "black", # we like the black borders of the points
             position = position_jitter(width = 0.3)) + #adding jitter as some points overlap
  geom_smooth(aes(color = site), #adding line to show how the relationship between carbs and proteins
              method = "lm", #using a linear regression for the trend line
              se = F, #removing the standard error 
              show.legend = F) + #don't show the legend
  theme_bw() + 
  labs(title = "Impact of Carbohydrates and Proteins on OTU Counts of Taxa Bacteroidia\nin Vietnamese and Tanzanian Latrine Pitts",
       x = "Protein",
       y = "Carbohydrates",
       caption = "Data from doi/10.1111/1751-7915.12334",
       fill = "",
       size = "OTU Count") +
  guides(color = guide_legend(override.aes = list(fill = NA))) + #remove the grey background from the legend
  theme(legend.position = "bottom") + #position the legend at the bottom
  scale_fill_brewer(palette = "Set1") + #change color of points
  scale_color_brewer(palette = "Set1") #Setting the palette to Set1 as this one is nicer
```

