---
title: "Lab 4 Assignment Group 4"
format: html
editor: visual
execute: 
  warning: false
---

### Group Exercise from Group 4

-   Edir Sebastian Vidal Castro (s243564),
-   Morten Orebo Holmström (s242223),
-   Nithiyashri Jayashankar (s244356),
-   Tilman Johannes Nelissen (s233043),
-   Thea Rehm Rosholm (s225019)

### Loading packages
```{r}
#| message: false
library("tidyverse")
library("patchwork")
library("here")
```



# Loading data

```{r}
diabetes_data <- read_csv(file = here("data/diabetes (1).csv"))
```



### T1
Change `height`, `weight`, `waist` and `hip` variables from imperical to metric system. We need to use `mutate` to do this but first we need to know how to recalcuate from imperical to metric. 
Conversion of inches to cm = inches x 2.54
Conversion of pounds to kg = pounds x 0.45359237
 
 And we should round to one decimal using the `round` function. 
 
```{r}
diabetes_data <- diabetes_data %>% 
  mutate(m_height = height * 2.54,
         m_weight = weight * 0.45359237,
         m_waist = waist * 2.54,
         m_hip = hip * 2.54) %>% 
           mutate(across(starts_with("m_"), 
                         ~round(., 1)))
         

```
 
Okay, so an important lesson here is that we can use the `mutate(across())` function to apply a subsequent function to mutate. 
So here we mutate across variables that `starts_with("m_")` which are our newly calculated variables, and then we apply the function `round` to these variables. As we use the `across` function we need to put the variables insinde round as a dot i.e. `round(., digits = 1)` and we need to provide a ~ before the `round` function. 


### Q17 
Creating new variable `BMI` which is `m_weight/((m_height/100)^2)`
```{r}
diabetes_data <- diabetes_data %>% 
  mutate(BMI = m_weight/((m_height/100)^2))
```


### T7 
Creating the variable `BFP` based on the equation 
BFP = 1.39 x BMI + 0.16 x age - 10 x gender - 9. 
where gender is 0 for female and 1 for male. 

So essentially we need to create a new column termed `gender_number` which gets 0 for female and 1 for male. 
```{r}
diabetes_data <- diabetes_data %>% 
  mutate(gender_number = case_when(gender == "female" ~ 0, 
                                   gender == "male" ~ 1))
```

And then we can construct our new variable based on the formula:
`mutate(BFP = 1.39 * BMI + 0.16 * age - 10 * gender_number - 9)`:
```{r}
diabetes_data <- diabetes_data %>% 
  mutate(BFP = 1.39 * BMI + 0.16 * age - 10 * gender_number - 9)
```


### T8 
Creating a `WHR`variable which is defined by: 
`mutate(WHR = m_waist/m_hip)`

```{r}
diabetes_data <- diabetes_data %>% 
  mutate(WHR = m_waist/m_hip)
```


### Q18
What correlates better with BMI - WHR or BFP? 

We can use the correlation coefficent for BMI vs. WHR and BFP respectively

BMI correlation to WHR:
```{r}
cor(diabetes_data$BMI, diabetes_data$WHR, use = "complete.obs")
```
Correlation coefficient between DMI and WHR of 0.1 - a low correlation coefficient. 


BMI correlation to BFP:
```{r}
cor(diabetes_data$BMI, diabetes_data$BFP, use = "complete.obs")
```
And for the BFP it is 0.89. Considerably higher correlation coefficient meaning, that BFP is better associated with BMI than WHR.  


We use the geom_point to asses the assocation between BMI, WHR and BFP graphically and save the plot in the object `whr_plot`. We apply a regression line to show the relationship between BMI and WHR
```{r}
whr_plot <- ggplot(data = diabetes_data, 
                   mapping = aes(x = BMI, 
                                 y = WHR)
                   ) + 
  geom_point(alpha =0.6,
             fill = "steelblue", 
             shape = 21, 
             size = 2) +
  labs(title = "Correlation between body mass\n index and waist to hip ratio", 
       y = "Waist to hip ratio", 
       x = "Body mass index",
       caption = "Pearsson correlation coefficent = 0.10")  + 
  geom_smooth(method = "lm",
              se = F) + 
  theme_bw()


whr_plot
```


We use the geom_point to asses the assocation between BMI, WHR and BFP graphically and save the plot in the object `bfp_plot`. We apply a regression line to show the relationship between BMI and BFP.
```{r}
bfp_plot <- ggplot(data = diabetes_data, 
                   mapping = aes(x = BMI, 
                                 y = BFP)
                   ) + 
  geom_point(alpha =0.6,
             fill = "steelblue", 
             shape = 21, 
             size = 2) +
  labs(title = "Correlation between body mass\n index and body fat percentage", 
       y = "Body fat percentage", 
       x = "Body mass index", 
       caption = "Pearsson correlation coefficent = 0.89")  + 
  geom_smooth(method = "lm",
              se = F) + 
  theme_bw()
bfp_plot
```

We use that `patchwork` package merge the two plots into one nice plot. 
```{r}
whr_plot + bfp_plot
```
